Bandit Multiclass Classification

### 2018-12-27 arxiv-version/main.tex
minor intro

% LaTeX packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[singlelinecheck=false]{caption}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{thmtools}
\usepackage{tikz}

@@

% Math symbols and commands
\newcommand{\R}{\mathbb{R}}  % set of real numbers
\newcommand{\indicator}[1]{\mathbf{1}\left[#1 \right]} % indicator
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle} % inner product
\newcommand{\norm}[1]{\left\| #1 \right\|}  % norm of a vector or a matrix
%\DeclareMathOperator*{\deg}{deg}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Exp}{\mathbf{E}}  % expected value
\DeclareMathOperator{\diff}{d \!} % differential

@@

\begin{abstract}
TODO
\end{abstract}

@@

We design new algorithms for the problem of \textsc{Online Multiclass
Classification with Bandit Feedback}. In this problem, in each round, the
algorithm receives a feature vector and the algorithm has to predict one of $K$
possible classes. Immediately after the algorithm makes its prediction, it
receives feedback whether the prediction was correct or not. However, in
contrast with the standard, so called full information setting, the algorithm
does \emph{not} receive the correct label. The feedback is only binary. To allow
for meaningful use of randomization, we assume that the correct label in a
particular round is chosen prior to algorithm's prediction in that
round.\footnote{All our results apply to also to a simpler model of
the so called \emph{oblivious adversary} where the whole sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is chosen before the first
round of the game.} The protocol of the problem is stated below.

\begin{algorithm}[h]
\caption{\textsc{Online Multiclass Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$. Number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t$}
\STATE{Adversary \emph{secretly} chooses correct class label $y_t \in \{1,2,\dots,K\}$}
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$}
\STATE{Observe feedback $z_t \in \{0,1\}$ where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

The goal is to design an algorithm that makes small number of mistakes
$\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$. The problem
was initially proposed by \cite{Kakade-Shalev-Shwartz-Tewari-2008}. The
corresponding problem with full information feedback is solved by the classical
\textsc{Multiclass Perceptron} algorithm and its variants (see e.g.
\cite{Crammer-Dekel-Keshet-Shalev-Shwartz-Singer-2006}). The bandit feedback
version is a mathematically natural model and potential applications include
online advertising and online content optimization where the binary feedback
corresponds to a click.

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}. In the full-information feedback
setting, it is well know that if the examples have norm at most $R$ and are
linearly separable with a margin $\gamma$ then the \textsc{Multiclass
Perceptron} algorithm makes at most $\lfloor 2(R/\gamma)^2 \rfloor$ mistakes.
This result is very satisfying since the upper bound on the number of mistakes
is information-theoretically optimal and at the same time the \textsc{Multiclass
Perceptron} algorithm has low time and memory complexity.

>>>>

% LaTeX packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[singlelinecheck=false]{caption}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{thmtools}
\usepackage{tikz}
\usepackage{color}
\usepackage{comment}

@@

% Math symbols and commands
\newcommand{\R}{\mathbb{R}}  % set of real numbers
\newcommand{\indicator}[1]{\mathbf{1}\left[#1 \right]} % indicator
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle} % inner product
\newcommand{\norm}[1]{\left\| #1 \right\|}  % norm of a vector or a matrix
%\DeclareMathOperator*{\deg}{deg}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Exp}{\mathbf{E}}  % expected value
\DeclareMathOperator{\diff}{d \!} % differential
\DeclareMathOperator*{\polylog}{polylog}

\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkred}{rgb}{0.7,0,0}
\definecolor{teal}{rgb}{0.3,0.8,0.8}
\definecolor{orange}{rgb}{1.0,0.5,0.0}
\definecolor{purple}{rgb}{0.8,0.0,0.8}
\definecolor{blue}{rgb}{0.0,0.0,1.0}
\newcommand{\kibitz}[2]{{\textcolor{#1}{\textsf{\footnotesize #2}}}}
\newcommand{\alina}[1]{\kibitz{darkred}{[AB: #1]}}
\newcommand{\david}[1]{\kibitz{darkgreen}{[DP: #1]}}
\newcommand{\balazs}[1]{\kibitz{purple}{[BS: #1]}}
\newcommand{\deva}[1]{\kibitz{teal}{[DT: #1]}}
\newcommand{\chenyu}[1]{\kibitz{orange}{[CW: #1]}}
\newcommand{\chicheng}[1]{\kibitz{blue}{[CZ: #1]}}

@@

\begin{abstract}
We study the problem of efficient online multiclass linear classification with bandit feedback, where all examples belong to one of $K$ classes and lie in the $d$-dimensional Euclidean space.
Previous works have left open the challenge of designing efficient algorithms with finite mistake bounds
when the data is linearly separable by a margin $\gamma$.
In this work, we take a first step towards this problem.
We identify two notions of linear separability:
one-versus-all (OvA) linear separability and weakly linear separability.
\begin{enumerate}
\item Under the OvA
linear separability condition, we design an efficient algorithm that achieves a near-optimal mistake bound of $O(\frac{K}{\gamma^2})$.
\item Under the more challenging weakly linear separability condition, we design an efficient algorithm with a mistake bound of $2^{\tilde{O}(\min(K \log^2 \frac 1 \gamma, \log K \sqrt{\frac 1 \gamma})) }$ \footnote{We use the notation $\tilde{O}(f(\cdot)) = O( f(\cdot) \polylog(f(\cdot)) )$.}.
Our algorithm is based on a reduction from weakly linear separability to OvA linear separability via an infinite dimensional monomial feature map, which in turn is inspired by ~\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}

%We focus on the setting where there are $K$ classes, and there is a $d$-dimensional multiclass linear classifier that separates all the examples shown with a positive margin $\gamma$.
%$2^{O(\min(K \log^2 K \log^2 \frac 1 \gamma, \log K \log \log K \sqrt \frac 1 \gamma \log \frac 1\gamma))}$
%$\min( (\log K \sqrt{\frac 1 \gamma})^{O(\log K \sqrt{\frac 1 \gamma})}, K^{O(K \log K \log^2 \frac1\gamma)} )$.

\end{abstract}

@@

We study the problem of \textsc{Online Multiclass Linear
Classification with Bandit Feedback}~\cite{Kakade-Shalev-Shwartz-Tewari-2008}.
The problem can be viewed as a game between the learner and the environment:
at each timestep $t$, the environment chooses an example $(x_t, y_t)$, where its feature
part $x_t \in \R^d$ is revealed to the learner.
\footnote{All our results apply to also to a simpler model of
the so called \emph{oblivious adversary} where the whole sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is chosen before the first
round of the game.}
Upon receiving this example, the
learner is asked to make a prediction $\hat{y}_t$.
the learner receives feedback immediately after making its prediction.
In contrast with the standard full information setting (where the feedback given is the correct label $y_t$), the feedback given is only a binary indicator
whether the prediction was correct or not. The protocol of the problem is formally stated below.

%In this problem, in each round, the
%learner receives a feature vector and is asked to predict one of $K$
%possible classes. To allow
%for meaningful use of randomization, we assume that the correct label in a
%particular round is chosen prior to algorithm's prediction in that
%round.
%Immediately after the learner makes its prediction, it
%receives feedback whether the prediction was correct or not. However, in
%contrast with the standard, so called full information setting, the algorithm
%does \emph{not} receive the correct label. The feedback is only binary.

\begin{algorithm}[h]
\caption{\textsc{Online Multiclass Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$. Number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t$}
\STATE{Adversary \emph{secretly} chooses correct class label $y_t \in \{1,2,\dots,K\}$}
\chicheng{Or we simply say: Adversary chooses example $(x_t, y_t)$, where $x_t$ is reveal to the learner.}
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$}
\STATE{Observe feedback $z_t \in \{0,1\}$ where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

The performance is of the learner is measured by its cumulative number of mistakes
$\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$. The bandit feedback
 is a natural feedback model motivated by applications including
online advertising and online content optimization, where the learner only gets to
see the loss of its action taken.
%where the binary feedback corresponds to a click
%The
%corresponding problem with full information feedback is solved by the classical
%\textsc{Multiclass Perceptron} algorithm and its variants (see e.g.
%\cite{Crammer-Dekel-Keshet-Shalev-Shwartz-Singer-2006}).

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}.
\chicheng{I wonder if we should introduce the two notions of linear separability here,
because strong linear separability is also worth mentioning in the full-information setting.
}

In the full-information feedback
setting, it is well known that if the examples have norm at most $R$ and are
linearly separable with a margin $\gamma$ then the \textsc{Multiclass
Perceptron} algorithm makes at most $\lfloor 2(R/\gamma)^2 \rfloor$ mistakes.
This result is very satisfying since the upper bound on the number of mistakes
is information-theoretically optimal and at the same time the \textsc{Multiclass
Perceptron} algorithm has low time and memory complexity.

### 2018-12-27 ptf_margin.tex
minor intro

\item $\phi_0(x) = ( \sqrt{ (\frac 1 2)^{i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1, \ldots, i_d \geq 0}$. This corresponds to the kernel $K(x,x') = \frac{1}{1 - \frac 1 2 \inner{x,x'}}$.
Note that for all $x$ such that $\| x \|_2 \leq 1$,
$\inner{\phi_0(x), \phi_0(x)} = K(x,x) \leq 2$.
\end{enumerate}

>>>>

\item $\phi_0(x) = ( \sqrt{ (\frac 1 2)^{i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1, \ldots, i_d \geq 0}$. This corresponds to the kernel $K(x,x') = \frac{1}{1 - \frac 1 2 \inner{x,x'}}$.
Note that for all $x$ such that $\| x \|_2 \leq 1$,
$\inner{\phi_0(x), \phi_0(x)} = K(x,x) \leq 2$.

\item $\phi(x) = \sqrt{{i_1+\ldots+i_d \choose i_1 \ldots i_d}} x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq s, i_1, \ldots, i_d \geq 0}$. This corresponds to the multinomial kernel $K(x,x') = \sum_{i=0}^s \inner{x,x'}^i$. Note that for all $x$ such that $\| x \|_2 \leq 1$,
$\inner{\phi(x), \phi(x)} = K(x,x) = \sum_{i=0}^s \|x\|^{2i} \leq s+1$.

\item $\phi(x) = (\sqrt{\frac{1}{i_1! \ldots i_d!}} x_1^{i_1} \ldots x_d^{i_d})_{i_1, \ldots, i_d \geq 0}$. This corresponds to the kernel $K(x,x') = \exp(\inner{x,y})$. Note that for all $x$ such that $\|x\|_2 \leq 1$, $\inner{\phi(x), \phi(x)} \leq e$.

\cz{Actually I wasn't able to get the RBF kernel working in our problem, as every term in the feature map will be a monomial term *with a $\exp(-\frac{\|x\|_2^2}{2})$ scaling*; it is unclear how to represent a polynomial as $\inner{c, \phi(x)}$.}

\end{enumerate}
